Assessing spatial correlations in pulsar timing arrays with the noise-marginalized optimal statistic

### 2017-12-20 manuscript/manuscript.tex
Trying to set up automated paper generation with travis
Trying travis
Editing wording in introduction

@@

Low-frequency gravitational waves (GWs) with frequencies of 
$10^{-9} - 10^{-7} \; \mathrm{Hz}$ can be observed with pulsar timing arrays (PTAs) 
composed of millisecond pulsars (MSPs) \cite{hd1983,fb1990}. 
The dominant astrophysical source in this frequency range is an isotropic stochastic 
gravitational wave background (GWB) 
made up of the incoherent superposition of GWs from inspiraling 
supermassive black hole binaries (SMBHBs) 
\citep{1995ApJ...446..543R, 2003ApJ...583..616J, 2003ApJ...590..691W}. 
By monitoring the ultra-regular periodic emission from these pulsars using radio telescopes, 
we can probe the dynamics of the spacetime through which the pulses travel. 
This is done by searching for correlations in the pulsar timing residuals, 
which measure the difference between the expected and observed pulse times of arrival. 
Current upper-limits on the stochastic background from PTAs are approaching 
theoretical predictions for the stochastic GWB \citep{PPTA2013,EPTA2015,NANOGrav9yr}

The primary data analysis technique used by PTAs is Bayesian inference 
to compare the probabilities of various models for the pulsar timing residuals, 
including one where the residuals contain evidence of an isotropic GW stochastic background 
\citep{vlm+2009,lah+2013}. 
Bayesian inference is a powerful technique because it allows us to compare arbitrarily complicated models. 
However, running a full Bayesian analysis is computationally intensive, 
particularly when searching for evidence of dipole spatial correlations -- 
``the smoking gun'' of a stochastic GWB.

The significance of a stochastic GWB can also be assessed using the 
optimal statistic, a frequentist estimator for the GWB amplitude \citep{abc+2009,demorest+2013,ccs+2015}. 
It is an important complement to Bayesian detection techniques 
because it provides an independent detection procedure, 
and it is significantly faster to compute. 
In particular, different spatial correlations can be tested with the optimal statistic 
in less than a second, 
whereas a Bayesian analysis with spatial correlations requires 
several weeks to run on a supercomputing cluster.

However, previous work has found that the optimal statistic gives biased results 
for PTAs with significant red noise in the individual pulsars due to the 
strong covariance between the individual red noise parameters 
and the GWB amplitude \sv{has this ever been discussed in a paper?}. 
In this paper, we present a technique for improving the accuracy of the optimal statistic 
by marginalizing the optimal statistic over the individual red noise parameters 
using the posterior distributions from a full Bayesian analysis. 
This hybrid approach produces a more accurate 
estimate of \Agw\ and its uncertainty. 
This method is more computationally intensive than previous methods 
for computing the optimal statistic, but still requires only minutes. 
Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
can be used to compute the optimal statistic with any choice of the ORF, 
allowing additional spatial correlations to be tested very quickly.

>>>>

Test travis...

@@

Long-wavelength gravitational waves (GWs) with frequencies of 
$10^{-9} - 10^{-7} \; \mathrm{Hz}$ can be observed with pulsar timing arrays (PTAs) 
composed of millisecond pulsars (MSPs) \cite{hd1983,fb1990}. 
The dominant astrophysical source in this frequency range is an isotropic stochastic 
gravitational wave background (GWB) 
made up of the incoherent superposition of GWs from inspiraling 
supermassive black hole binaries (SMBHBs) 
\citep{1995ApJ...446..543R, 2003ApJ...583..616J, 2003ApJ...590..691W}. 
By monitoring the ultra-regular periodic emission from these pulsars using radio telescopes, 
we can probe the dynamics of the spacetime through which the pulses travel. 
This is done by searching for correlations in the pulsar timing residuals, 
which measure the difference between the expected and observed pulse times of arrival. 
Current upper-limits on the stochastic background from PTAs are approaching 
theoretical predictions for the stochastic GWB \citep{PPTA2013,EPTA2015,NANOGrav9yr}

The primary data analysis technique used by PTAs is Bayesian inference 
to compare the probabilities of various models for the pulsar timing residuals, 
including one where the residuals contain evidence of an isotropic GW stochastic background 
\citep{vlm+2009,lah+2013}. 
Bayesian inference is a powerful technique because it allows us to compare arbitrarily complicated models. 
However, running a full Bayesian analysis is computationally intensive, 
particularly when searching for evidence of dipole spatial correlations:
``the smoking gun'' of a stochastic GWB.

The significance of a stochastic GWB can also be assessed using the 
optimal statistic, a frequentist estimator for the GWB amplitude \citep{abc+2009,demorest+2013,ccs+2015}. 
%It is an important complement to Bayesian detection techniques 
%because it provides an independent detection procedure, 
%and it is significantly faster to compute. 
%In particular, different spatial correlations can be tested with the optimal statistic 
%in less than a second, 
%whereas a Bayesian analysis with spatial correlations requires 
%several weeks to run on a supercomputing cluster.
Not only does it provide an independent detection procedure, complementing a more robust Bayesian analysis, but its calculation requires significantly less time to compute. In particular, the optimal statistic produces results for various spatial correlation functions, a bottleneck within the more traditional stochastic GWB analysis, within seconds; a full Bayesian analysis including correlations has to run for several weeks on a supercomputing cluster.

However, previous work has found that the optimal statistic gives biased results 
for PTAs with significant red noise in the individual pulsars due to the 
strong covariance between the individual red noise parameters 
and the GWB amplitude \sv{has this ever been discussed in a paper?}. 
%In this paper, we present a technique for improving the accuracy of the optimal statistic 
%by marginalizing the optimal statistic over the individual red noise parameters 
%using the posterior distributions from a full Bayesian analysis. 
%This hybrid approach produces a more accurate 
%estimate of \Agw\ and its uncertainty. 
Here we present a technique for improving the accuracy of the optimal statistic by including an additional step: marginalizing over the individual pulsar red noise parameters using the posterior distributions from a full PTA Bayesian analysis. This hybrid approach produces a more precise estimate of the stochastic GWB amplitude, 
\Agw\/, and its uncertainty, while requiring only a few minutes over the more traditional method of computation. 
%This method is more computationally intensive than previous methods 
%for computing the optimal statistic, but still requires only minutes. 
%Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
%can be used to compute the optimal statistic with any choice of the ORF, 
%allowing additional spatial correlations to be tested very quickly.
Furthermore, the same Bayesian analysis drawn upon by the noise marginalization can be used to compute the optimal statistic with any choice of overlap-reduction function (ORF), allowing for quick testing of various spatial correlations.
